\part{Log}

\chapter{Onboarding}

\section*{First Week 10/1- 10/7}
Begin date: 10/2

\begin{itemize}
\item Refamiliarized myself with Shell scripting. Notes are described in my local computational science directory.
\item Went over notes of objectives from yesterday's ApTSi meeting
\end{itemize}


Begin date: 10/3

The first thing I did was watch a few YouTube videos on 2x speed related to Apache Kafka. Through that, I found out that I'd need Java, Spark, and Zookeeper. I also found out that I'd have the easiest time using a Unix-like system such as WSL Ubuntu.

\subsubsection*{Install Java}
Attempted to install Java development kit (JDK) from \href{https://www.oracle.com/java/technologies/javase-jdk11-downloads.html}{Oracle}.

Installation was easier on Linux.
\begin{itemize}
\item At the shell, use \verb|java --version| to see if it's already installed.
\item It will prompt installation with \verb+apt install openjdk-11-jre-headless+ or something similar. After that installs, \verb|java --version| should work.
\end{itemize}


Begin date: 10/4

Attempted to install Kafka w/ quickstart and didn't have enough of the pre-requisite technologies.

\subsubsection*{Install Docker}
From \href{https://docs.docker.com/engine/install/ubuntu/}{docs.docker.com/.../ubuntu/}:

\begin{itemize}
\item To uninstall old versions:
\begin{lstlisting}[language=bash]
sudo apt-get remove docker docker-engine docker.io containerd runc
\end{lstlisting}

\item Before installing Docker Engine for the first time on a new host machine, you need to set up the Docker repository. Afterward, the Docker can be installed and updated form the repository.

Update \texttt{apt} package index and install packaes to allow \texttt{apt} to use a repository over HTTPS:
\begin{lstlisting}[language=bash]
sudo apt-get install \
apt-transport-https \
ca-certificates \
curl \
gnupg-agent \
software-properties-common
\end{lstlisting}

Add Docker's official GPG key:
\begin{lstlisting}[language=bash]
curl -fsSL https://download.docker.com/linux/ubuntu/gpg | sudo apt-key add -
\end{lstlisting}

To verify you have the key:
\begin{lstlisting}
sudo apt-key fingerprint 0EBFCD88
\end{lstlisting}

Continue the rest of the installation guide until running the hello world script.
\end{itemize}

I'm getting a lot of errors when trying to run the "hello world" command for Docker. I'm currently trying a bunch of suggesstions from forums and stack exchange

Begin date: 10/6

\subsubsection*{Install Zookeeper}




\section*{Week 2, 10/8- 10/14}
date: 10/8 -10/9

\subsection*{Docker Concepts}






