\chapter{Miscellaneous}


\section{HTTP, REST, \& Spring Boot}

\begin{quest}
\item \cloze
REST has become the de-facto standard for building web services. Why? Easy to build, easy to consume.

\item
What is a micro service and why do we use them?
\begin{ans}
Microservice is an architecture that allows the developers to develop and deploy services independently. Each service running has its own process and this achieves the lightweight model to support business applications.

Microservices, a.k.a. the microservice architecture, structure an application as a collection of services that are
	\begin{itemize}
	\item Highly maintainable and testable
	\item Loosely coupled
	\item Independently deployable
	\item Organized around business capabilities
	\item Owned by a small team
	\end{itemize}
\end{ans}

\item What is Spring Boot?
\begin{ans}
An open source Java-based framework used to create microservices.
\end{ans}

\item Why use Spring Boot?
\begin{ans}
	\begin{itemize}
	\item manages REST endpoints
	\item flexible way to configure database transactions
	\item auto configured as opposed to manually configured
	\item eases dependency management
	\end{itemize}
\end{ans}

\item How does Spring Boot easy dependency management?
\begin{ans}
Spring Boot automatically configures your application based on the dependencies you have added to the project by using @EnableAutoConfiguration annotation. For example, if MySQL database is on your classpath, but you have not configured any database connection, then Spring Boot auto-configures an in-memory database.

The entry point of the spring boot application is the class contains @SpringBootApplication annotation and the main method.
\end{ans}
\end{quest}





\subsection{References \& Further Reading}

\href{https://spring.io/guides/tutorials/rest/}{Building REST services with Spring.}

\href{https://microservices.io/#:~:text=Microservices%20%2D%20also%20known%20as%20the,Organized%20around%20business%20capabilities}{What are microservices?}

\href{https://www.tutorialspoint.com/spring_boot/spring_boot_introduction.htm}{Spring Boot - Introduction}