\chapter{Algorithms}

% -----------------------------------------------------
\section{Algorithms}
% -----------------------------------------------------


\begin{quote}
	"4.5 years of learning programming and working as fullstack software engineer ... had interview with one of the FAANG companies this summer in Hong Kong but failed it due to the fact that I suck in DSA (Data Structures \& Algorithms)."
\end{quote}

\begin{quote}
	"I'm using to leetcode.com to learn data structures
	and algorithms since I got a rejection from FAANG after interviewing with them onsite."
\end{quote}

\href{https://blog.codechef.com/2020/07/24/the-role-of-data-structure-and-algorithms-in-programming/}{Role of DSA in Programming (July, 2020)}





% -----------------------------------------------------
\section{Design Patterns}
% -----------------------------------------------------

\begin{quest}
\item Why use "design patterns"?
\begin{ans}
\begin{itemize}
	\item Design patterns let your write better code more quickly by providing a clearer picture of how to implement the design
	\item Design patterns encourage code reuse and accomodate change by supplying well-tested mechanisms for delegation, composition, and other non-inheritance based reuse techniques
	\item Design patterns encourage more legible and maintainable code
\end{itemize}
\end{ans}

\item Delegation? Composition?
\begin{ans}
\begin{itemize}
	\item delegation: a pattern where a given object provides an interface to a set of operations. However, the actual work for those operations is performed by one or more other objects.
	\item composition: Creating objects with other objects as members. Should be used when a "has-a" relationship appears.
\end{itemize}
\end{ans}

\item What are design patterns?
\begin{ans}
	content

\end{ans}

\item Which resources will you use to start learning about design patterns?
\begin{ans}
	GOF patterns (C++). Then, potentially Head First Design Patterns (Java)/
\end{ans}
\end{quest}

\subsection{References \& Further Reading}

\href{https://www.gofpatterns.com/design-patterns/module1/intro-design-patterns.php}{Introduction to Design Patterns Course}
