
\chapter{Daily Scratch}
\section{Aug 9-15}

\subsection*{Who is François Chollet?}
\begin{itemize}
\item
	works on deep learning at Google in Mountain View, CA.
\item
	creator of the Keras deep-learning library
\item
	contributor to the TensorFlow machine-learning framework.
\item
	wrote a book on deep learning that is very applied rather than theoretical.
\end{itemize}

\begin{quote}
``We know the field is fast moving. If the reader looking for more recent free reading resources, there are some good introductory/tutorial/survey papers on Arxiv; I happen to be compiling a list of them''. - Auto Snipe
\end{quote}

 One of said review papers \cite{raghu2020survey}


\hrule
\section{Medical Research Meetings}

\subsection*{Aug 18 | Marwah}


\subsection*{Aug 24 |  Itsif}

\hrule
\section{Algo Trading}

\subsection{ML Finance Project}

\paragraph*{Terms to Learn: } NLP, bag of words, stop words, stemming, tokenizing, term-frequency inverse document frequency, bigrams, unigrams, Dow Jones Industrial Average (DJIA) index

\begin{quotation}
	`` The preprocessing for the text data involved three parts: removing stop words, stemming, and tokenizing. I first removed common words like "the" and "a". This also included removing non-standard english words, such as emojis and words in other languages. The next step was performing stemming, which turns words like "running" to "run". From there, I tried many methods of tokenizing the corpus, including bag of words, term frequency–inverse document frequency, and using bigrams and unigrams. For this particular dataset, a bag of words model using unigrams worked the best.

	Using our data from the bag of words model, I merged each day with whether or not the Dow Jones Industrial Average (DJIA) index will move up or down for the next day. I thought the movement of the DJIA would able to capture the health of the overall economy while not giving too much information into how tech stocks are specifically doing...

	The top words associated with DJIA moving up are "gadafi", "Iran" and "sanctions". ''
\end{quotation}


\subsection{\href{https://stackabuse.com/time-series-prediction-using-lstm-with-pytorch-in-python/}{Time Series in Pytorch | Introductory Example}}

\begin{quest}
	\item
	\cloze
	Neural networks can be constructed using the \pyth{torch.nn} package.

	\item
	Import the package for constructing neural networks in PyTorch.
	\begin{ans}
		\pyth{import torch.nn as nn}
	\end{ans}

	\item \cloze Seaborn comes with built-in datasets.

	\item Load seaborn's flights dataset.
	\begin{ans}
		\pyth{flight_data = sns.load_dataset("flights")}
	\end{ans}

	\item
	Why must time series data be scaled for sequence predictions?
	\begin{ans}
		When a network is fit on unscaled data, it is possible for large inputs to slow down the learning and convergence of your network and in some cases prevent the network from effectively learning your problem.
	\end{ans}

	\item
	sklearn import for scaling data?
	\begin{ans}
		\pyth{from sklearn.preprocessing import MinMaxScaler}
	\end{ans}
\end{quest}

\section{Daily Schedule (when free)}

Lex Fridman posted a video today (2020 Aug 27), detailing  \href{https://youtu.be/0m3hGZvD-0s}{his daily} schedule outside of work commitments. The stages are as follows:

\begin{tabular}{lllll}
	(1) Morning & (2) Mantra  &  (3) Deep work I & (4) Medium depth  & (5) Exercise \\
	(6) Shower & (7) Deep work II & (8) Eat & (9) Shallow work  & (10) Paper reading \\
\end{tabular}

I'll go into a bit more depth on the sections below and try to implement some of these practices in my daily schedule to see what I can learn from Lex.

\begin{enumerate}
\item
	Morning

\item
	Mantra / Daily Review
	\begin{itemize}
	\item
		Rules \& constraints
	\item
		Gratitude
	\item
		Long-term goals
	\item
		Short-term goals
	\item
		Core principles
	\end{itemize}

\item
	Deep Work I (4 hr)

\item
	Social Media and Music Practice

\item
	Exercise

	Exercise every day, no matter what. Even if injured, find a body part and exercise it.

\item
	Shower

\item
	Deep Work II (4 hr)

\item
	Eat

	2k calories, all your nutrients
	vegetables, protein, and fat

\item
	Shallow Work (1 hr)

\item
	Paper Reading (1 hr)
\end{enumerate}

\section{Software Project Portfolio}
\paragraph*{Key factors}
\begin{itemize}
\item Completeness
\item Functionality
\item Readability
\item Documentation/information

\end{itemize}

\section{Hugo}

Start Sep 16 (Mike Dane Tutorial Series)

\subsection{Intro to Hugo -\href{https://youtu.be/qtIqKaDlqXo}{(video)} }
\begin{itemize}
\item
	Hugo is a static site generator.
\item
	Static website generators allow you to compromise between writing a bunch of static html pages and using a heavy, and potentially expensive, content management system.
\item
	Why Hugo? It's extremely fast.
\item
	2 kinds of websites, dynamic and static. Dyanmic ex. Facebook. Facebook pages are dynamically generated for each user. For static websites, what you see is what you get.
\item
	Static websites are notoriously harder to maintain b/c you lack some fo the flexibility of things  on a dynamic site. Usually you can't use much conditional logic, functions, or variables.
\item
	However, static pages are extremely fast.
\item
	Hugo is great for a blog, portfolio website, etc.
\item
	Hugo doesn't explicity require you to write a single line of HTML code.
\item
	Flexibility | With that said, if you want to go in and change every little detail of the layout of the site, you cna do that. You can write as much of the HTML and have as much control as you'd like.
\item
	Hugo is 100\% free and open-source.
\end{itemize}

\subsection{Intalling Hugo on Windows - \href{https://youtu.be/G7umPCU-8xc?list=PLLAZ4kZ9dFpOnyRlyS-liKL5ReHDcj4G3}{(video)} }
\begin{itemize}
\item
	Mine's already installed. I'll skip this for now.
\item
	a
\item
	a
\end{itemize}

\subsection{Creating a new site - \href{https://youtu.be/sB0HLHjgQ7E?list=PLLAZ4kZ9dFpOnyRlyS-liKL5ReHDcj4G3}{(video)} }
\begin{itemize}
\item skip, a bit too easy

\end{itemize}

\subsection{Installing \& Using Themes - \href{https://youtu.be/L34JL_3Jkyc?list=PLLAZ4kZ9dFpOnyRlyS-liKL5ReHDcj4G3}{(video)} }
\begin{itemize}
\item my themes are already installed
\end{itemize}

\subsection{Creating \& Organizing Content - \href{https://youtu.be/0GZxidrlaRM?list=PLLAZ4kZ9dFpOnyRlyS-liKL5ReHDcj4G3}{(video)} }
\begin{itemize}
\item
	Hugo has 2 types of content: single pages and list pages
\item
	List content lists other content on the site. You can call this a list page.
\item
	Individual blog posts are single pages.
\item
	Your posts should not just be in the content directory. They should be in directories inside the content directory.
\item
	A list page is automatically created for directories inside the content folder.  Hugo automatically does this. Note, this only occurs for directories at the ``root" level of the content directory. For example:  \texttt{content/post/} would generate a list page at \texttt{site.com/post/}, but \texttt{content/post/dir0} would not.
\item
	If you want a list page to be generated for a dir that is not at the root level of the content dir, you have to create an \textbf{index filed}, \texttt{\_index.md}. For a convenient and efficient way to do this from the cmd, use \texttt{hugo new post/dir0/\_index.md} (above above example), then there will be a list page for dir0. 	Content can also be added to\texttt{\_index.md} and it should show up on the page.
\item
	Additionally, for list pages that are aautomatically generate by hugo, you can edit the content by adding an index.md to those as well. Ex. \texttt{~content/post/\_index.md}.
\end{itemize}

\subsection{Front Matter- \href{https://youtu.be/Yh2xKRJGff4?list=PLLAZ4kZ9dFpOnyRlyS-liKL5ReHDcj4G3 }{(video)} }
\begin{itemize}
\item
	Front matter in Hugo is what is commonly called meta data.
\item
	Front matter is data about our content files.
\item
	The metadata automatically generated by Hugo at the top of md files when using \texttt{hugo new } is front matter
\item
	Front matter is stored in key-value pairs
\item
	Front matter can be written in 3 different languages: YAML, TOML, and JSON
\item
	The defualt lang for front matter in Hugo is YAML
\item
	YAML - indicated by ``--'',
\item
	TOML - indicated by "+++" and uses "=" instead of ":",
\item
	JSON - indicated
\item
	You can create your own custom front matter variables.
\item
	Front matter is super powerful in its utility.
\end{itemize}


\subsection{Archetypes - \href{https://youtu.be/bcme8AzVh6o?list=PLLAZ4kZ9dFpOnyRlyS-liKL5ReHDcj4G3 }{(video)} }
\begin{itemize}
\item
	How does the default front matter from using \texttt{hugo new ~.md} get selected? Short answer: archetypes
\item
	An archetype is basically the default front matter template for when you create a new content file.
\item
	Archetypes are modified under \texttt{static/themes/archetypes/default.md}
\item
	Suppose your content dir has a subdirectory, \texttt{content/dir0}. If you wanted to create an archetype for the files in dir0, you'd simply create \texttt{dir0.md} inside the archetypes dir.
\end{itemize}

\subsection{Shortcodes - \href{https://youtu.be/2xkNJL4gJ9E?list=PLLAZ4kZ9dFpOnyRlyS-liKL5ReHDcj4G3 }{(video)} }
\begin{itemize}
\item
	Shortcodes are predefined chunks of HTML that you can insert into your markdown files.
\item
	Let's say you have a md file that you want to spice up by adding in some custom HTML. For instance, maybe you'd like to embed a YouTube video. Normally this would require lots of HTML that you'd have to paste it. Shortcodes can allow you to sidestep this. Hugo comes with a YouTube video shortcode predefined.
\item
	 General shortcode syntax \texttt{{{< shortcode\_name param0 >}}}
\item
	Youtube shortcode | For a YouTube video with url, ``youtube.com/watch?v=random-text", the shortcode we'd use to embed would be \texttt{{{< youtube random-text >}}} because ``random-text'' is the id of the youtube video and the only parameter for that shortcode.
\end{itemize}

\subsection{Taxonomies - \href{https://youtu.be/pCPCQgqC8RA?list=PLLAZ4kZ9dFpOnyRlyS-liKL5ReHDcj4G3 }{(video)}}
\begin{itemize}
\item
	Taxonomies in hugo are basically ways that you can logically group different pieces of content together in order to organize it in a more efficient way.
\item
	Hugo provids 2 defualt taxonomies: tags \& categories
\item
	All taxonomy information is declared in front matter.  In YAML, tags has the syntax \texttt{tags: ["tag0", "tag1", $\ldots$]}
\end{itemize}

\subsection{Templates - \href{https://youtu.be/gnJbPO-GFIw}{(video)}}
\begin{itemize}
\item
	Templates here mostly refers to HTML templates. If you're not comfortable writing HTML, CSS, and coding for the web, templates might be a little bit above your head.
\item
	A hugo theme is actually made up of hugo templates.
\item
	Any template that you use in Hugo is going to be inside \texttt{themes/theme-name/layouts}. This is where all the templates live.
\item
	\texttt{~/layouts/default} usually contains a default style for list and single pages by use of \texttt{list.html} and \texttt{single.html}.
\end{itemize}

\subsection{List Templates - \href{https://youtu.be/8b2YTSMdMps}{(video)}}
\begin{itemize}
\item
	List templates give default HTML layout to list content files.
\item

\end{itemize}

\subsection*{Resources}
\begin{itemize}
\item
	\href{https://www.linkedin.com/learning/learning-static-site-building-with-hugo-2/build-a-static-site-with-hugo?resume=false}{A clear and concise beginner hugo tutorial}
\item
	\href{https://youtu.be/yfoY53QXEnI}{CSS Crash Course for Absolute Beginners}
\end{itemize}

